### 2017-04-02 chapters/introduction/patient.tex
Add latest version

As discussed in \ref{sec:mech_boudary}, we constrain the basal movement in
the longitudinal direction, which is easiest to accomplished using a flat
base located at a prescribed location. In order to make the base flat
we first orient the surfaces so that the longitudinal axis is aligned
with the $x$-axis, and the apex pointing in the postive $x$
direction. To conststruct the basal plane we first take out the basal
points from the strain mesh, and use these points to construct a least
sqauare fitting plane (Figure \ref{fig:strain_mesh_plane}). Let $(x_i, y_i, z_i), i = 1, \cdots, N$
be the basal points, and suppose the basal plane solves the equation
$z = ax + by + c$, for some unknown constants $a,b,c$. Following a
least square approach, we select the parameters $(a,b,c)$ that
minimizes to sum
\begin{align}
  \sum_{i = 1}^{N} \left( z_i - ax_i + by_i + c \right)^2.
\end{align}
Once the parameters for the basal plane is found we adjust the size of
the cut, by moving the plane along the longitudinal axis (here
$x-$axis), until the cavity volume agrees with the measured volume
given by the image segmentation tool within some specified tolerance. 
When the correct size is found the points above the plane is removed,
and the remaining surfaces are smoothed using the GAMer
\cite{yu2008feature}. The actualy mesh generation is performed using Gmsh
\cite{geuzaine2009gmsh}, which meshes the endocardial and epicardial
surface togther unsing frontal-Delaunay meshing algorithm. Gmsh also marks
the endocardal, the epicardial and the basal facets so that it is easy
to

@@

@@

Dolfin-Adjoint is a software package that based on the FEniCS project
which aims to derrive the discrete adjoint equation and the tangent
linear models of finite element models implemented within the FEniCS
framework. While the tradional approach is to derrive the adjoint code
from the forward code, Dolfin-Adjoint utilises the high level symbolic
representation \cite{UFL} in FEniCS to derrive the discrete adjoint equations
from the discrete forward eqations, and then the FEniCS system
derrives the adjoint code from the discrete adjoint equations.

@@

\subsection{Identifiability of parameters}
Try dividing heart mesh is different segment, discuss uniqueness of
paraperers
Discuss regularization technices

>>>>

As discussed in \ref{sec:mech_boudary}, we constrain the basal movement in
the longitudinal direction, which is easiest to accomplished using a flat
base located at a prescribed location. In order to make the base flat
we first orient the surfaces so that the longitudinal axis is aligned
with the $x$-axis, and the apex pointing in the postive $x$
direction. To conststruct the basal plane we first take out the basal
points from the strain mesh, and use these points to construct a least
sqauare fitting plane (Figure \ref{fig:strain_mesh_plane}). Let $(x_i, y_i, z_i), i = 1, \cdots, N$
be the basal points, and suppose the basal plane solves the equation
$z = ax + by + c$, for some unknown constants $a,b,c$. Following a
least square approach, we select the parameters $(a,b,c)$ that
minimizes to sum
\begin{align}
  \sum_{i = 1}^{N} \left( z_i - ax_i + by_i + c \right)^2.
\end{align}
Once the parameters for the basal plane is found we adjust the size of
the cut, by moving the plane along the longitudinal axis (here
$x-$axis), until the cavity volume agrees with the measured volume
given by the image segmentation tool within some specified tolerance. 
When the correct size is found the points above the plane is removed,
and the remaining surfaces are smoothed using the GAMer
\cite{yu2008feature}.

The actualy mesh generation is performed using Gmsh
\cite{geuzaine2009gmsh}, which meshes the endocardial and epicardial
surface togther unsing frontal-Delaunay meshing algorithm. Gmsh also marks
the endocardal, the epicardial  and the basal facets, along with the
endocardial

\begin{figure}[htbp]
  \centering
  \begin{subfigure}[t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{chapters/introduction/figures/geometry/mesh.png}
    \caption{\label{fig:mesh}}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{chapters/introduction/figures/geometry/markers_full.pdf}
    \caption{\label{fig:markers_full}}
  \end{subfigure}
\caption{}
\label{fig:echopac_output}
\end{figure}

@@

\subsubsection{Rule-based fibers}
\label{sec:rule_based_fiber}
Mention papers about fiber distribution, fiber dependence and
different rule-based fibers.

Describe the main steps in the Bayer algorithm.

\subsubsection{The ventricular coordinate system}

When dealing with geometric shapes such as a left ventricle, finding a
coordinate system that is easy to work with is important. For
example, when studing spherical shapes, the spherical coordinate system is
usually easier to work with. Likewise, a coordinate system that is
easy to use when studying ellipsoidal shapes like the left ventricle is
the \emph{prolate spheroidal coordinate system} \cite{hunter1996kd}. 
The cartesian coordinates $(x,y,z)$ are related to the prolate
spheroidal coordinates by

\begin{align}
  \begin{split}
    x &= a \sinh \mu \sin \nu \cos \theta  \\
    y &= a \sinh \mu \sin \nu \sin \theta  \\
    z &= a \cosh \mu \cos \nu,
  \end{split}
\end{align}
where $a$ is the focal point, $\nu \in [0, \pi]$ is the longitudinal
coordinate, $\theta \in [0, 2\pi]$ is the azimuthal angle and $\mu$
is the radial coordinate. For an ellipse given by the equation
$\frac{x^2}{b^2} +\frac{y^2}{c^2}  = 1$ with $b$ being the semimajor
axis, and $c$ the semiminor axis, the focal point is given by
\begin{align}
  a =\sqrt{b^2 - c^2}.
  \label{eq:focal_point}
\end{align}
Assuming the ventricle is axisymetric around the
longitudinal axis, which is aligned with the $x$-axis it is possible
to estimate the focal point by taking $b$ in \eqref{eq:focal_point} to
be the maximum distance from the base to the apex, and $b$ the maximum
radius at the based. Using this estimate, the base is assumed to be
located approximately at the center of the ellipsoid, i.e $\nu = \pi /
2$. Compared to geometries used in benchmarking of cardiac mechanics
problems, this is not a totally wrong assumption
\cite{land2015verification}

\begin{figure}[htbp]
  \centering
  \begin{subfigure}[t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{chapters/introduction/figures/geometry/prolate.png}
    \caption{\label{fig:prolate_coord}}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \includegraphics[width=\textwidth]{chapters/introduction/figures/geometry/prolate_cube.png}
    \caption{\label{fig:prolate_cube}}
  \end{subfigure}
\caption{}
\label{fig:prolate}
\end{figure}

\subsubsection{Marking the mesh according to the AHA-segments}

@@

Dolfin-Adjoint is a software package that based on the FEniCS project
which aims to derrive the discrete adjoint equation and the tangent
linear models of finite element models implemented within the FEniCS
framework. While the tradional approach is to derrive the adjoint code
from the forward code using automatic differentiation tools,
Dolfin-Adjoint utilises the high level symbolic representation
\cite{UFL} in FEniCS to derrive the discrete adjoint equations from
the discrete forward eqations, and then the FEniCS system derrives the
adjoint code from the discrete adjoint equations.

@@

Line search methods, interior point methods, gradient based methods,
genetic methods, trust region

\subsection{Identifiability of parameters}
% Try dividing heart mesh is different segment, discuss uniqueness of
% paraperers
% Discuss regularization technices
When dealing with parameter estimation, you should always consider
questions about identifiability and uniquness. This depends on the
model, the data and the objective functional
\cite{hadjicharalambous2015analysis}. A first test should be to
generate synthetic data with some prescribed parameters, feed this
error-free data to the data assimilation method, and ensure that you
are able to retrieve the parameters that generated that data. If you
end up with a different parameter set than the one generated the data,
your problem is not \emph{structuraly identifiable}
\cite{chabiniok2016multiphysics}. Your problem is said to be
\emph{practical identifiable} if the parameters can be determined
uniquely based on the data at hand. If the test passes you shold try
to add noise to the data, and see if your data assimilation is stable
with respect to noise in the measurement, and you should quantify the
model out error versus the model input error. Demonstrating identifiability is
often difficult, especially when your data is corrupted with noise,
and when the complexity of the model increases. Balancing the
complexity of model in terms of \emph{model fidelity}, i.e wheter your
model is rich enough to representing the data, and the identifiability
is often key in order to have robust model. A too simple model will
often fail to represent your data, while a too complex model, might
give completely differet answers with just a slight perturbation of
the data, i.e with noise added to the measurement. The same problem is
encountered in other scientific disiplines, for instance in
statistical learning theory, \emph{overfitting} is related to
overparameterization of the model, in which you can perfectly
represent your data, but a small perturbation will cause a significant
error. Another problem, which is also related to identifiability is
the question about uniquness of the soluton. A practical test is to
start the optimization algorithm from different intial points and see
if you end up at the same optimum. If this does not happen, it is
likely that the objective functional is not convex, which is a
requirement for uniquness. Unfortunately, this is the case for many
problems in PDE-constrained optimization. In these cases, the objective
functional might have several distinct local minima, in which case
different techniques for find the best local minima exists
\cite{farrell2015multiple}.

\subsubsection{An example of a non-structuraly identifiable problem}
Split the ventricle in two, and use volume as the only input.
Show that differnt parameter sets (with $\gamma$) will generate the
same volume

\subsubsection{Material parameter estimation}

\paragraph{One volume measurement only}

\paragraph{Multiple volume measurements}

\paragraph{Volume and strain measurements}

\subsubsection{Coupled unloading and material parameter estimation}
In Section \ref{sef:reference_geometry} we described a method for
finding the uloaded configuration unsing the backward displacement
method. The problem of using this method when dealing with parameter
estimation is that the unloaded geometry depends on the material
properties, which in general is unknown. Hence we need a way of
estimating both the unloaded geometry and the material parameters at
the same time. One possibible algorithm is the following:

\begin{algorithm}
\caption{Coupled Unloading and material parameters estimation}\label{alg:unloaded_material}
\begin{algorithmic}[1]
  \State $i = 0$
  \State residual = $\infty$
  \State $\mvec_0 =  \mvec$ \Comment{Initial guess for material parameters}
  \While{i $<$ N and residual > tolerance }

  \State $\Omega_{\mathcal{U}}^i$ = Unload($\Omega_{\mathcal{I}}, \mvec_i$) \Comment{unload
    material}
  \State $\mvec_{i+1}$ = EstimateMaterial($\Omega_{\mathcal{U}}^i,
  \mvec_i$) \Comment{estimate material parameters}
  \State residual = ComputeResidual($\Omega_{\mathcal{U}}^i,
  \Omega_{\mathcal{U}}^{i-1}$)
  \State $i = i +1$

  \EndWhile
\end{algorithmic}
\end{algorithm}

\subsection{Activation parameter estimation}

\subsubsection{Multiobjetive optimization}
Paramaters in the model are estimated based on minimizing a cost
functional representing the mismatch between observations and simulaton.
Assume we are given $N$ different observations, then in priciple we
have $N$ different cost funcitonals to solve
\begin{equation}
  \begin{aligned}
    \label{eq:multiobjetive_opt}
    & \text{minimize}
    & &  \left\{ \mathcal{J}_1(\lambda), \mathcal{J}_2(\lambda), \cdots, \mathcal{J}_N(\lambda) \right\} \\
    & \text{subject to}
    & & \delta \Pi(\state(\lambda)) = 0, \lambda \in \mathcal{P},
  \end{aligned}
\end{equation}
whith 
\begin{align}
  \mathcal{J}_i(\lambda) = \left( \mathbf{y} - \mathcal{H}(\state(\lambda)) \right)^2, \; i = 1,\cdots, N
\end{align}
The above observations comes from the same soure, and hence it should
be intuitive that minimizing one of them, would also bring the other
closer to a minimum. However, with precense of noise in the data,
this might not always be the case, and hence we have a conflicting
objectives. Problem \eqref{eq:multiobjetive_opt} is called a
multiobjetive optimization problem \cite{deb2016multi}. Such problems
typically do not have a single optimal solution, but a family of
solutions called \emph{Pareto optmal solutions}. The basic methods for
solving such problems inclues \emph{the weighted method}, \emph{the
  $\varepsilon$-constraint method},

### 2017-04-02 chapters/introduction/physiology.tex
Add latest version

Deoxgenated blood is transported from the body trough the veins and to
the right atrium (RA). From the RA, blood flows to the right ventricle
(RV), before it is sent through the the pulmonary ateries(PA) and to the
lungs, where the bloods gets oxygenated. From the lungs, the blood
travels trough the pulmonary veins to the left atrium (LA), and
finally to the left ventricle (LV), before is is sent out to the body
again trough the aorta (Ao). A schamtic view of the blood flow
through the body is shown in Figure ~\ref{}

>>>>

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.8\textwidth]{chapters/introduction/figures/heart_anatomy.pdf}
\caption{Backward displacement method might fail for bi-ventricular
  geometries. This can be seen be noticing that the RV free wall is
  buckling and colliding with the septum.}
\label{fig:heart_anatomy}
\end{figure}

Deoxgenated blood is transported from the body trough the veins and to
the right atrium (RA). From the RA, blood flows to the right ventricle
(RV), before it is sent through the the pulmonary ateries(PA) and to the
lungs, where the bloods gets oxygenated. From the lungs, the blood
travels trough the pulmonary veins to the left atrium (LA), and
finally to the left ventricle (LV), before is is sent out to the body
again trough the aorta (Ao). A schamtic view of the blood flow
through the body is shown in Figure ~\ref{fig:heart_anatomy}
